---
title: "Biodiversity Analysis"
author: "Andrea Martinez"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      fig.path = "../figures/04_Biodiversity/") # send any figure output to this folder
```

# Setting the Environment

## Set the seed
```{r set-seed}
set.seed(238428)
```

## Load Libraries
```{r load-packages}
pacman:: p_load(tidyverse, devtools, patchwork, iNEXT, phyloseq,
                install = FALSE)
```

## Load in Data
```{r load-data}
load("/local/workdir/alm379/git_repos/SalinityGradient_16S/analysis/02_PreProcessing/raw_preprocessed_phyloseq.RData")
raw_preprocessed_phyloseq

# Intuition check
min(sample_sums(raw_preprocessed_phyloseq))

# Make metadata dataframe
metadata_df<-raw_preprocessed_phyloseq%>%
  sample_data()%>%
  data.frame()

# Setting colors for the stations
station_colors <- c(
  "Shipping Channel" = "dodgerblue4",
  "Aransas Bay" = "dodgerblue2",
  "Copano West" = "#D9CC3C",
  "Copano East" = "#A0E0BA",
  "Mesquite Bay" = "#00ADA7")
```

# Goals
1. Calculate the Hill Diversity of the samples.
2. Evaluate the rarefaction curves.
3. Evaluate the diversity values
4. Make notes of specific samples and their seq depth

# Diversity Calculations with iNEXT

```{r calc-div}
#prepare input data
iNEXT_input_df <- raw_preprocessed_phyloseq %>%
  otu_table()%>%
  data.frame()
# Quick check
dim(iNEXT_input_df)

# Run iNEXT: Calculate the Hill Numbers
# Note that: Species in ROWs, samples in COLUMNS
# Remember to set the seed!
#iNEXT_data <- iNEXT(iNEXT_input_df,
#                    q = c(0,1,2), datatype = "abundance")
# Save the file
#save(iNEXT_data, file = "data/04_Biodiversity/iNEXT_data.RData")
```

# Evaluate the Diversity!
```{r load-div}
load("data/04_Biodiversity/iNEXT_data.RData")
str(iNEXT_data)
typeof(iNEXT_data)
```

# Plot Diversity
```{r plot-rarefaction}
# Prepare Colors

color_df <-
  iNEXT_input_df %>%
  colnames() %>%
  data.frame()
# Rename the column
colnames(color_df)[1] <- "names"
# Check
head(color_df)

# Make a helper dataframe for plotting with colors
iNEXT_color_df <- color_df%>%
  # Fix the names for merging 
  mutate(names = gsub(names, pattern = "[.]", replace = "-"), 
         names = gsub(names, pattern = "X", replace = ""))%>%
  # Merge with metadata
  left_join(metadata_df, by = "names")%>%
  # Merge with colors for plotting with ggiNEXT
  left_join(data.frame(station_colors = station_colors,
           station = names(station_colors)),
           by = "station")
```

#Plot Rarefaction with 'ggiNEXT'

```{r ggiNEXT, fig.width=8, fig.height =3.5}

# Plot rarefaction!
# Rarefaction/ectrapolation cure, type =1

# Order q:
# 0 = Richness/total number ot taxa
# 1 = Exponential Shannon / Number of "Common" taxa
# 2 = Inverse Simpson / Number of "Dominant" taxa

ggiNEXT(iNEXT_data, type =1, facet.var = "Order.q") +
  facet_wrap(~Order.q, scales = "fixed")+
  scale_color_manual(values = iNEXT_color_df$station_colors, guide = FALSE)+
  scale_fill_manual(values = iNEXT_color_df$station_colors, guide = FALSE)+
  scale_shape_manual(values = base::rep(17, nsamples(raw_preprocessed_phyloseq)),
                     guide = FALSE)+
    theme(legend.position = "none")
```

# Manually plot Diversity

## Rarefaction
```{r iNEXT-manual}
iNEXT_manual_df <- iNEXT_data$iNextEst$size_based %>%
  dplyr::rename(names = Assemblage) %>%
  # Fix the sample names
  mutate(names = gsub(names, pattern = "[.]", replace = "-"),
         names = gsub(names, pattern = "X", replace = "")) %>%
  # join with metadata
left_join(., metadata_df, by = "names") %>%
  # add colors to data frame
left_join(., data.frame(station_colors = station_colors,
          station = names(station_colors)),
by = "station")

dim(iNEXT_manual_df)
str(iNEXT_manual_df)

# Plot it - Rarefaction Curve
iNEXT_manual_df %>%
  # filter out rows that are calculated by rarefaction from iNEXT
  dplyr::filter(Method == "Extrapolation") %>% 
  # Make the actual plot
  # the # of sequences on the x-axis and diversity on the y-axis
  # You can choose to pick one diversity value or plot all three
  ggplot(aes(x=m , y=qD, color = station, group = names, group_by = names)) +
  geom_line() +
  # geom_point() +
  facet_grid(Order.q ~ station, scales = "fixed") +
  scale_color_manual(values = station_colors) +
  theme(legend.position = "bottom")
```


# Environmetal variable change
```{r environmental-pairs, fig.width =8, fig.height =8}
head(metadata_df)
# Pull out environmental variables

env_df <-
  metadata_df %>%
  dplyr::select(names, water_tempC:DO_mgL)

# inspect
head(env_df)

# Plot the correlations
pairs(dplyr::select(env_df, -names), upper.pNEL = NULL)

```


#Diversity versus salinity
```{r div-vs-salinity, fig.width = 3.5, fig.height = 6}

#PSU = practical salinity unit
# 1 PSU = 1 g of salt per 1,000 grams of water
# 1 PSU = 1 ppt (part per thousand)
# Freshwater = 0-0.5
# Brackish = 0.5-30
# Marine = >30

iNEXT_manual_df %>%
  dplyr::filter(Method == "Observed") %>%
  ggplot(aes(x=salinity_psu, y =qD))+
  facet_wrap(.~Order.q, scales = "free")+
  geom_point(aes(color = station))+
  stat_smooth(method = "lm", formula = y ~poly(x, 2))+
  labs(x = "Salinity (PSU)", y = "# of ASVs")+
  scale_color_manual(values = station_colors)+
  theme(legend.position = "bottom")
```


#Session Information
```{r session-info}
devtools:: session_info()
```

